Fast Sorting on Modern Computers

### 2005-08-14 docs/tech_report/elementary.tex
fixed the captions for the graphs, and a few of the graphs themselves

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32-byte cache line and separate instruction cache.]
{\label{elementary level 1 misses}\includegraphics[scale=\myscale]{plots/on2_-_Level_1_misses.eps}}
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}.]
{\label{elementary branches}\includegraphics[scale=\myscale]{plots/on2_-_Branches.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Cache and branch prediction simulation results for elementary sorts}
\label{Cache and branch prediction simulation results for elementary sorts}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\centering
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred}, with bimodal and two-level adaptive
predictors. The simulated two-level adaptive predictor used a 10 bit branch history register which was XOR-ed with the
program counter.]
{\includegraphics[scale=\myscale]{plots/on2_-_Branch_misses1.eps}}
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred}, with bimodal and two-level adaptive
predictors. The simulated two-level adaptive predictor used a 10 bit branch history register which was XOR-ed with the
program counter.]
{\includegraphics[scale=\myscale]{plots/on2_-_Branch_misses2.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Branch simulation results for elementary sorts}
\label{Branch simulation results for elementary sorts}
\end{figure}
}

>>>>

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar
\cc{sim-cache}, simulating an 8KB data cache with a 32-byte cache line and
separate instruction cache. Results shown are for a direct mapped cache.]
{\label{elementary level 1 misses}\includegraphics[scale=\myscale]{plots/on2_-_Level_1_misses.eps}}
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}.]
{\label{elementary branches}\includegraphics[scale=\myscale]{plots/on2_-_Branches.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Cache and branch prediction simulation results for elementary sorts}
\label{Cache and branch prediction simulation results for elementary sorts}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\centering
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred},
with bimodal and two-level adaptive predictors. The simulated two-level adaptive
predictor used a 10 bit branch history register which was XOR-ed with the
program counter. Results shown are for a 512 byte bimodal predictor.]
{\includegraphics[scale=\myscale]{plots/on2_-_Branch_misses1.eps}}
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred},
with bimodal and two-level adaptive predictors. The simulated two-level adaptive
predictor used a 10 bit branch history register which was XOR-ed with the
program counter. Results shown are for a 512 byte bimodal predictor.]
{\includegraphics[scale=\myscale]{plots/on2_-_Branch_misses2.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Branch simulation results for elementary sorts}
\label{Branch simulation results for elementary sorts}
\end{figure}
}

### 2005-08-14 docs/tech_report/report.tex
fixed captitalization in captions
fixed the captions for the graphs, and a few of the graphs themselves

\newenvironment{changemargin}{%
 \begin{list}{}{%
\setlength{\textwidth}{\paperwidth}
\setlength{\textheight}{\paperheight}
%\setlength{\oddsidemargin}{-1in}
%\setlength{\evensidemargin}{-1in}
\setlength{\topmargin}{-1in}
\setlength{\topsep}{-1in}%
	\setlength{\leftmargin}{-1.5in}%
	\setlength{\rightmargin}{-1.5in}%
	\setlength{\listparindent}{\parindent}%
	\setlength{\itemindent}{\parindent}%
	\setlength{\parsep}{\parskip}%
 }%
\centering%
\item[]%
}{\end{list}}

@@

%cc is used for
% - unix commans or programs: gcc, sim-bpred, sim-cache, espresso, li, simple,
%   sim-fast, diff, valgrind, cachegrind, driver
% - code or keywords: less, Item, exch, if, while, unsigned int, int, qsort,
%   goto, do-while, THRESHOLD, partition, malloc, shellsort, improved shellsort
% - file names: main.c
% - #defines: UNIT_MAX
\newcommand{\cc}[1]{\texttt{#1}}

%n is used for
% - proper names: Atom, SimpleScalar, Papi, GnuPlot, ValGrind, glibc
% - technical names: unsigned integer, less-than, greater-than, in-place, out-of
%   place, Abstract Data Type, records, cache hierarchy, latency, hit, miss,
%   compulsory, cold-start, capacity, conflict, cache line, cache block, offset,
%   tag, virtaul addresses, bus address, temporal locality, locality, spatial
%   locality, taken, not taken, flow control branch, comparaitve branch, static,
%   semi-static, dynamic, 1-bit, bimodal, two-level adaptive, heap, children,
%   implicit heap, heapifying, sifting, fix-up, fix-down, sift-up, sift-down,
%   Repeated-adds, Remove-min, d-heap, fanout, natural, straight, unrolled,
%   Memory management unit, tiling, k-way merge, trashing divide and conquer,
%   multi-way partition, comparision based, counting based, increments
% - web addresses: http://www.gnuplot.com, http://valgrind.kde.org
% - sort names, when highlighted: algorithm S, algorithm N
% - branch predictor names: 'Creation, child promotion', etc
\newcommand{\n}[1]{\textit{#1}}

\newcommand{\code}[2]{\begin{figure}\lstinputlisting{code/#2}\caption{#1}\label{#1}\end{figure}}

\newcommand{\myscale}{1.2}
\newcommand{\plot}[7]{
% 3 figures first, then 3 page for them
\afterpage{
\setcounter{subfigure}{0}

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key - this was measured on a Pentium 4 using hardware performance counters. #2]
{\label{#1 cycles}\includegraphics[scale=\myscale]{plots/#1_-_cycles.eps}}
\subfigure[Instructions per key - this was simulated using SimpleScalar \cc{sim-cache}. #3]	
{\label{#1 instructions}\includegraphics[scale=\myscale]{plots/#1_-_Instructions.eps}}
\end{changemargin}
\vspace*{7em}
\caption{Simulated Instruction count and empiric cycle count for #1sort}
\label{Simulated Instruction count and empiric cycle count for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32-byte cache line and separate instruction cache. #4]
{\label{#1 level 1 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_1_misses.eps}}\\
\subfigure[Level 2 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating a 2MB data
and instruction cache and a 32-byte cache line. #5]
{\label{#1 level 2 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_2_misses.eps}}\\
\end{changemargin}
\vspace*{7em}
\caption{Cache Simulation results for #1sort}
\label{Cache Simulation results for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-7em}
\begin{figure}[H]
\begin{changemargin}
\centering
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}. #6]
{\label{#1 branches}\includegraphics[scale=\myscale]{plots/#1_-_Branches.eps}}
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred}, with bimodal and two-level adaptive
predictors. The simulated two-level adaptive predictor used a 10 bit branch history register which was XOR-ed with the
program counter. #7]
{\label{#1 branch misses}\includegraphics[scale=\myscale]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Branch Simulation results for #1sort}
\label{Branch Simulation results for #1sort}
\end{figure}
}}

\newcommand{\smallscale}{.74}
\newcommand{\smallplot}[2]{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key] {\includegraphics[scale=0.74]{plots/#1-_cycles.eps}}
\subfigure[Instructions per key] {\includegraphics[scale=0.74]{plots/#1_-_Instructions.eps}}
\subfigure[Level 1 cache misses per key]	{\includegraphics[scale=0.74]{plots/#1_-_Level_1_misses.eps}}
\subfigure[Level 2 cache misses per key]	{\label{#2 level 2 misses}\includegraphics[scale=0.74]{plots/#1_-_Level_2_misses.eps}}
\subfigure[Branches per key]				{\includegraphics[scale=0.74]{plots/#1_-_Branches.eps}}
\subfigure[Branch misses per key]			{\includegraphics[scale=0.74]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Simulation results for #2}
\label{Simulation results for #2}
\end{figure}
\newpage
}

\newcommand{\sixplot}[7]{
\afterpage{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[]{\label{#7-1}\includegraphics[scale=0.74]{plots/counter_#1}}
\ifthenelse{\equal{#2}{}}{}{\subfigure[]{\label{#7-2}\includegraphics[scale=0.74]{plots/counter_#2}}}
\ifthenelse{\equal{#3}{}}{}{\subfigure[]{\label{#7-3}\includegraphics[scale=0.74]{plots/counter_#3}}}
\ifthenelse{\equal{#4}{}}{}{\subfigure[]{\label{#7-4}\includegraphics[scale=0.74]{plots/counter_#4}}}
\ifthenelse{\equal{#5}{}}{}{\subfigure[]{\label{#7-5}\includegraphics[scale=0.74]{plots/counter_#5}}}
\ifthenelse{\equal{#6}{}}{}{\subfigure[]{\label{#7-6}\includegraphics[scale=0.74]{plots/counter_#6}}}
\end{changemargin}
\vspace*{6em}
\caption[Branch prediction performance for #7]{Branch prediction performance for #7. All graphs use the same scale.}
\label{Branch prediction performance for #7}
\end{figure}
\newpage
}}

\lstset{language=c,tabsize=3,basicstyle=\small,showspaces=false,showtabs=false,showstringspaces=false,stepnumber=1,numberstyle=\small,frame=tblr}

\long\def\symbolfootnote[#1]#2{\begingroup\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

@@

\hyphenation{La-Marca}

@@

% TODO the bugs are now fixed (actually, one of them was only mistakenly a bg :)
% the only remaining thing is the compulsory cache subtraction.
\chapter{Bug List}
\input bugs.tex

>>>>

% widens the margins for the graphs
\newenvironment{changemargin}{%
 \begin{list}{}{%
\setlength{\textwidth}{\paperwidth}
\setlength{\textheight}{\paperheight}
%\setlength{\oddsidemargin}{-1in}
%\setlength{\evensidemargin}{-1in}
\setlength{\topmargin}{-1in}
\setlength{\topsep}{-1in}%
	\setlength{\leftmargin}{-1.5in}%
	\setlength{\rightmargin}{-1.5in}%
	\setlength{\listparindent}{\parindent}%
	\setlength{\itemindent}{\parindent}%
	\setlength{\parsep}{\parskip}%
 }%
\centering%
\item[]%
}{\end{list}}

@@

%\cc{x} is used for
% - unix commans or programs: gcc, sim-bpred, sim-cache, espresso, li, simple,
%   sim-fast, diff, valgrind, cachegrind, driver
% - code or keywords: less, Item, exch, if, while, unsigned int, int, qsort,
%   goto, do-while, THRESHOLD, partition, malloc, shellsort, improved shellsort
% - file names: main.c
% - #defines: UNIT_MAX
\newcommand{\cc}[1]{\texttt{#1}}

%\n{x} is used for
% - proper names: Atom, SimpleScalar, Papi, GnuPlot, ValGrind, glibc
% - technical names: unsigned integer, less-than, greater-than, in-place, out-of
%   place, Abstract Data Type, records, cache hierarchy, latency, hit, miss,
%   compulsory, cold-start, capacity, conflict, cache line, cache block, offset,
%   tag, virtaul addresses, bus address, temporal locality, locality, spatial
%   locality, taken, not taken, flow control branch, comparaitve branch, static,
%   semi-static, dynamic, 1-bit, bimodal, two-level adaptive, heap, children,
%   implicit heap, heapifying, sifting, fix-up, fix-down, sift-up, sift-down,
%   Repeated-adds, Remove-min, d-heap, fanout, natural, straight, unrolled,
%   Memory management unit, tiling, k-way merge, trashing divide and conquer,
%   multi-way partition, comparision based, counting based, increments
% - web addresses: http://www.gnuplot.com, http://valgrind.kde.org
% - sort names, when highlighted: algorithm S, algorithm N
% - branch predictor names: 'Creation, child promotion', etc
\newcommand{\n}[1]{\textit{#1}}

% to include a file as code
\newcommand{\code}[2]{\begin{figure}\lstinputlisting{code/#2}\caption{#1}\label{#1}\end{figure}}

% used for scaling graphs
\newcommand{\myscale}{1.2}

% plots graphs across 3 pages
\newcommand{\plot}[7]{
% 3 figures first, then 3 page for them
\afterpage{
\setcounter{subfigure}{0}

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key - this was measured on a Pentium 4 using hardware performance counters. #2]
{\label{#1 cycles}\includegraphics[scale=\myscale]{plots/#1_-_cycles.eps}}
\subfigure[Instructions per key - this was simulated using SimpleScalar \cc{sim-cache}. #3]	
{\label{#1 instructions}\includegraphics[scale=\myscale]{plots/#1_-_Instructions.eps}}
\end{changemargin}
\vspace*{7em}
\caption{Simulated instruction count and empiric cycle count for #1sort}
\label{Simulated instruction count and empiric cycle count for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar
\cc{sim-cache}, simulating an 8KB data cache with a 32-byte cache line and
separate instruction cache. #4] {\label{#1 level 1
misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_1_misses.eps}}\\
\subfigure[Level 2 cache misses per key - this was simulated using SimpleScalar
\cc{sim-cache}, simulating a 2MB data and instruction cache and a 32-byte cache
line. #5]
{\label{#1 level 2 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_2_misses.eps}}\\
\end{changemargin}
\vspace*{7em}
\caption{Cache simulation results for #1sort}
\label{Cache simulation results for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-7em}
\begin{figure}[H]
\begin{changemargin}
\centering
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}. #6]
{\label{#1 branches}\includegraphics[scale=\myscale]{plots/#1_-_Branches.eps}}
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred},
with bimodal and two-level adaptive predictors. The simulated two-level adaptive
predictor used a 10-bit branch history register which was XOR-ed with the
program counter. #7]
{\label{#1 branch misses}\includegraphics[scale=\myscale]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Branch simulation results for #1sort}
\label{Branch simulation results for #1sort}
\end{figure}
}}

% the scaling for the smaller graphs
\newcommand{\smallscale}{.74}

% plots 6 images on 1 page
\newcommand{\smallplot}[2]{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key] {\includegraphics[scale=0.74]{plots/#1-_cycles.eps}}
\subfigure[Instructions per key] {\includegraphics[scale=0.74]{plots/#1_-_Instructions.eps}}
\subfigure[Level 1 cache misses per key]	{\includegraphics[scale=0.74]{plots/#1_-_Level_1_misses.eps}}
\subfigure[Level 2 cache misses per key]	{\label{#2 level 2 misses}\includegraphics[scale=0.74]{plots/#1_-_Level_2_misses.eps}}
\subfigure[Branches per key]				{\includegraphics[scale=0.74]{plots/#1_-_Branches.eps}}
\subfigure[Branch misses per key]			{\includegraphics[scale=0.74]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Simulation results for #2}
\label{Simulation results for #2}
\end{figure}
\newpage
}

% plots 6 or less images on a page, using the counters
\newcommand{\sixplot}[7]{
\afterpage{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[]{\label{#7-1}\includegraphics[scale=0.74]{plots/counter_#1}}
\ifthenelse{\equal{#2}{}}{}{\subfigure[]{\label{#7-2}\includegraphics[scale=0.74]{plots/counter_#2}}}
\ifthenelse{\equal{#3}{}}{}{\subfigure[]{\label{#7-3}\includegraphics[scale=0.74]{plots/counter_#3}}}
\ifthenelse{\equal{#4}{}}{}{\subfigure[]{\label{#7-4}\includegraphics[scale=0.74]{plots/counter_#4}}}
\ifthenelse{\equal{#5}{}}{}{\subfigure[]{\label{#7-5}\includegraphics[scale=0.74]{plots/counter_#5}}}
\ifthenelse{\equal{#6}{}}{}{\subfigure[]{\label{#7-6}\includegraphics[scale=0.74]{plots/counter_#6}}}
\end{changemargin}
\vspace*{6em}
\caption[Branch prediction performance for #7]{Branch prediction performance for #7. All graphs use the same scale.}
\label{Branch prediction performance for #7}
\end{figure}
\newpage
}}

% settings for the code listings
\lstset{language=c,tabsize=3,basicstyle=\small,showspaces=false,showtabs=false,showstringspaces=false,stepnumber=1,numberstyle=\small,frame=tblr}

% i actually have no idea what this does, and no memory of writing it
\long\def\symbolfootnote[#1]#2{\begingroup\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

@@

% how to hyphenate words which arent standard
\hyphenation{La-Marca}

@@

% the only remaining thing is the compulsory cache subtraction.
\chapter{Bug List}
\input bugs.tex
